---
title: "Work_with_html_xml_json 7"
author: "anjal hussan"
date: "3/18/2018"
output: 
  html_document:
    theme: lumen
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective

Our goal is to create a list of books stored in an HTML table, an XML file, and a JSON file and use R to read and parse the files and store in 3 data frames.

# Environment Prep
```{r, warning=FALSE, message=FALSE}
if (!require('rvest')) install.packages('rvest')
if (!require('XML')) install.packages('XML')
if (!require('jsonlite')) install.packages('jsonlite')
library('DT')
```

# Read Data

## HTML Table
```{r}
h <- read_html("books.html") 
h.df <- data.frame(html_table(h))
knitr::kable(h.df)
```

## XML
```{r}
x <- xmlParse("books.xml")
x.df <- xmlToDataFrame(x)
knitr::kable(x.df)
```

## JSON
```{r}
j <- fromJSON("books.json")
j.df <- do.call("rbind", lapply(j, data.frame, stringsAsFactors=FALSE))
rownames(j.df) <- NULL
knitr::kable(j.df)
```

# Similarities

Overall, all the datafram looks same. But we need some deterministic test to confirm if they are same. We can use the base package function 'all.equal' to test.    

## HTML to XML
```{r}
all.equal(h.df, x.df)
```

## HTML to JSON
```{r}
all.equal(h.df, j.df)
```

## XML to JSON
```{r}
all.equal(x.df, j.df)
```

# Conclusion
Although the data frames look same, it would take some additional work to make them completely equivalent. The classes are different. Most importantly the xml import method mad each variable a factor.




