---
title: "Tidydata"
author: "anjal hussan"
date: "3/4/2018"
output: 
  html_document:
    theme: lumen
    toc: yes
---

## Objective
Our objective is to take a .csv of untidy data of two airlines and their on time or delayed flying information into multiple destination. Once the data is loaced, we need to transform and tidy the data, and then do an analysis to compare the arrival delays for both airlines.

## Setup

### Environment Prep

```{r, warning=FALSE, message=FALSE}
if (!require("dplyr")) install.packages('dplyr')
if (!require("tidyr")) install.packages('tidyr')
```

### Data Import

```{r}
airlines <- read.csv("https://raw.githubusercontent.com/ahussan/DATA_607_CUNY_SPS/master/Week5_Assignment/airlineData.csv", strip.white=TRUE)
knitr::kable(airlines)
```

## Data Transformation and Cleaning

```{r, comment=NA}
# remove middle NA obs
airlines <- airlines[-3,]

# add in missing airline values
airlines[2,1] <- "ALASKA"
airlines[4,1] <- "AMWEST"

# rename cols
names(airlines)[c(1,2)] <- c("Airline", "Status")

# tidy data
tidy <- airlines %>% 
    gather(cities, count, 3:7) %>% 
    spread(Status, count)

#update col
names(tidy)[4] <- "ontime"

# let's review our progress
knitr::kable(tidy)

```


## Data Analysis

### Ratios by City

```{r}
#Delay ratios
tidy <- tidy %>%
    mutate(total=ontime + delayed, delayratio=round(delayed/total, 3)) 

#Plot these delay ratios
dotchart(tidy$delayratio, groups = tidy$Airline, labels = tidy$cities , main = "Ratio of Delayed Flights by City")
```


### Ratios in Total
```{r}
summary <- tidy %>% 
    group_by(Airline) %>% 
    summarise(delayed = sum(delayed), ontime = sum(ontime), total = sum(total)) %>%
    mutate(delayratio = round(delayed/total, 3))

knitr::kable(summary)

#Plot delay ratios
dotchart(summary$delayratio, groups = summary$Airline, main = "Ratio of Delayed Flights Across All Cities")

```

## Conclusion

The individual city delay ratios by airline indicates that ALASKA airlines had a lower individual delay ratio than AMWEST in every single city. But the aggregated data shows that AMWEST had a lower overall delay ratio, due to the larger volume of flights.   
