---
title: "Project 1 - Chess Rating"
author: "anjal hussan"
date: "2/24/2018"
output: 
  html_document:
    theme: lumen
    toc: yes
---

### Load required package

```{r package, warning = FALSE, message=FALSE}
if (!require('DT')) install.packages('DT')
if (!require('stringr')) install.packages('stringr')
if (!require('ggplot2')) install.packages('ggplot2')
```

## load tournament information

```{r}
sourcefile = 'https://raw.githubusercontent.com/ahussan/DATA_607_CUNY_SPS/master/Project%201/tournamentinfo.txt'
raw <- readLines('tournamentinfo.txt')
head(raw)
```
### Data Grouping and Splitting
Our data is laid out in a table, it will be helpful to narrow down the areas our regular expressions to match the pattern

We may be able to extract necessary fields by matching pattern from certain sections of a line.

```{r, comment=NA}
#Find the table breaks
b0 <- 0
b1 <- unname(str_locate_all(pattern = '\\|', raw[5])[[1]][1,1])
b2 <- unname(str_locate_all(pattern = '\\|', raw[5])[[1]][2,1])
b3 <- unname(str_locate_all(pattern = '\\|', raw[5])[[1]][3,1])
b4 <- max(nchar(raw))
```

In the initial data we can find required data in the following string positions in each line: 

* Section 1 (Number, State): `r b0` - `r b1`
* Section 2 (Name, Pre-Rating): `r b1` - `r b2`
* Section 3 (Points): `r b2` - `r b3`
* Section 4 (Opponent IDs): `r b3` - `r b4`


### Data seperating

Now that we have some matched pattern - we will divide into two groups to get our row data in to to section

```{r, comment=NA}
# Group1 = Num, Name, Points, Opponent IDs
g1row <- seq(5, 196, 3)
# Group2 = State, Rating
g2row <- seq(6, 196, 3)
# subset for easy searching
group1 <- raw[g1row]
group2 <- raw[g2row]
```

## Data Field Creation and Extraction
Now we need to extract our desired fields. 


### Player Name 
```{r, comment=NA}
# Find subsets of player name
namesub <- substr(group1, b1+1, b2-2)
namesub <- str_trim(namesub)
PlayerName <- str_to_title(namesub)
head(PlayerName)
```

### Player State
```{r}
statesub <- substr(group2, b0, b1-1)
State <- str_trim(statesub)
head(State)
# dataframe
chess <- data.frame(PlayerName, State)
```

### Total Points
```{r}
pointsub <- substr(group1, b2+1, b3-1)
head(pointsub)
chess$TotalPoints <- sprintf("%.1f", as.numeric(pointsub))
```

### Player Pre-Rating

```{r}
presub <- substr(group2, b1+1, b2-1)
presub <- str_extract(presub, ': *\\d{2,}')
head(presub)
chess$PreRating <- as.integer(str_extract(presub, '\\d{2,}'))
```

### Avg Opponent Pre-Rating

```{r}
oppsub <- substr(group1, b3+1, b4)
oppsub <- str_extract_all(oppsub, '\\b\\d{1,}')
oppsub <- as.matrix(oppsub)

calculation <- function(l, p ){
  temp <- l[p]
    for (place in temp){
        rating <- 0
        counter <- 0
        for(i in place) {
            counter <- counter + 1
            rating <- rating + chess$PreRating[as.numeric(i)]
        }
        rating <- round(rating / counter)
    }
  return (rating)
}

chess$AvgOppPreRating <- apply(oppsub, 1, calculation)
```

## Final Dataset

### Review
```{r}
datatable(chess)
```

### Export in .csv

```{r}
# export
write.csv(chess, "chessData.csv", row.names=FALSE)
```

### Data visualization

```{r}
x <- ggplot(chess, aes(PreRating, AvgOppPreRating)) + geom_point(aes(color=TotalPoints)) + ggtitle("Pre-Rating VS Avg Opponent Pre-Rating by Total Points Gained")
x
```
