---
title: "Web APIs; Week 9"
author: "anjal hussan"
date: "4/8/2018"
output: 
  html_document:
    theme: lumen
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Our task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe. 

# Environment Prep
```{r, warning=FALSE, message=FALSE}
if (!require('jsonlite')) install.packages('jsonlite')
if (!require('dplyr')) install.packages('dplyr')
if (!require('stringr')) install.packages('stringr')
library('DT')
```

# Importing

We will be using Top Stories V1 API and look for stories in the Technology section.  

## Key Config

I stored API developer Key in a .R file and sourced accordingly in the code. This file won't be checked in to the git repository

```{r}
# get our API key
source("key.R")

# add text required for call
addkey <- paste0("&api-key=", key)
```

## GET

Get call to the API endpoint and get the feed

```{r}
# add url and text required for call
url <- "https://api.nytimes.com/svc/topstories/v1/technology.json"
addurl <- paste0(url, "?")

# make call
feed <- fromJSON(paste0(addurl, addkey))

```

# Setup DF

```{r}
sub <- feed$results
sub$pub_date <- unlist(str_extract(sub$published_date, ".+?(?=T)"))
sub$pub_time <- unlist(str_extract(sub$published_date, "(?<=T)(.*)"))

sub <- sub %>% 
    filter(section == "Technology") %>% 
    select(title, abstract, byline, url, pub_date, pub_time)
```

## Conclusion

```{r}

knitr::kable(sub)
```
