---
title: "Week3HW Regular Expressions"
author: "anjal hussan"
date: "2/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3. Extract Alpha RegEx

```{r, comment=NA, warning=FALSE}
library(stringr)

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"   

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.
(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
(c) Construct a logical vector indicating whether a character has a second name.

```{r, comment=NA}
### Function to extract last names
get_last <- function(list){
    last <- str_extract(list, '[[:alpha:]]{1,}\\,|\\b [[:alpha:]]{2,}')
    last <- str_extract(last, "[[:alpha:]]{1,}")
    last
}

### Function to extract first names
get_first <- function(list) {
    first <- str_extract(list, '[[:alpha:]]{1,} |\\. [[:alpha:]]{1,}|\\, [[:alpha:]]{2,}')
    first <- str_extract(first, "[[:alpha:]]{1,}")
    first
}

# Create df
allNameInDF <- data.frame(first = get_first(name), 
                      last = get_last(name))

allNameInDF$fullname <- paste0(allNameInDF$first, " ", allNameInDF$last)

# Find if title exists
allNameInDF$title <- str_detect(name, 'Dr.|Rev.')

# Find if Middle name exists
allNameInDF$MiddleName <- str_detect(name, ' [:alpha:]{1}\\. ')

# Print DF
allNameInDF
```

## 4. RegEx Construction

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

#### a)
This regex will return a string of one or more numbers followed by a '$' sign.  

```{r, comment=NA}
#(a) [0-9]+\\$
a <- "892$xp09$.2*.2$oa$"
unlist(str_extract_all(a, '[0-9]+\\$'))
```

#### b)
This regex will return a string of one to four lowercase alpha characters so long as they are bounded by non-word characters. Note that digits are considered to be word characters, thus the "two" does not show up because the "3" is not a boundary, and is not [a-z].   

```{r, comment=NA}
#(b) \b[a-z]{1,4}\\b
b <- "u-one 3two.three FOUR%four$;etc U pp Er"
unlist(str_extract_all(b, '\\b[a-z]{1,4}\\b'))
```

#### c)
This regex will return any string that ends with '.txt'.   If it does not end with .txt, it will not return anything.
```{r, comment=NA}
#(c) .*?\\.txt$
c <- "This$$is Ending with .txt"
c2 <- "this .txt is not ending with txt"
unlist(str_extract_all(c, '.*?\\.txt$'))
```

#### d) dd/dd/dddd
This regex will return any string in the form 'dd/dd/dddd' where 'd' is a digit.   This is a common form for dates.  
```{r, comment=NA}
#(d) \\d{2}/\\d{2}/\\d{4}
d <- "09/10/2016 10/12/2014 2015/01/02"
unlist(str_extract_all(d, '\\d{2}/\\d{2}/\\d{4}'))
```

#### e)
This regex returns any string that starts with a ``<text>`` and ends with ``</text>``. 
```{r, comment=NA}
#(e) <(.+?)>.+?</\\1>
e <- "<div>HTML tags </div> <ol><li>This is LI Tag</li><li>Again LI Tag</li></ol>"
unlist(str_extract_all(e, '<(.+?)>.+?</\\1>'))
```
